Malacology: A Programmable Storage System

### 2016-10-12 paper/paper.tex
paper: fix zlog sequencers with mantle evaluation - remove mds since its an implementation detail - explain the scenario - emphasize that we are testing the load balancing interface
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\item An implementation of the programmable Mantle metadata load balancing
service~\cite{sevilla:sc15-mantle}

@@

\begin{table*}
\centering
\title{Hi}
\begin{tabular}{  l | l | l | l }
\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Versioning}
  & Zookeeper coordination~\cite{hunt_zookeeper_2010}, Chubby~\cite{website:chubby}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization \& batching
  \\
{\it Logical Device}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
  & migration \& monitoring
  \\
\end{tabular}
\caption{The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces contributed in this paper.}
\label{table:examples}
\end{table*}

@@

{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
metadata load balancer, has been re-implemented on Malacology so that its
infrastructure can enjoy all the properties of other Ceph components. The load
balancing logic has stayed largely the same but the infrastructure has been
improved to safely load, manage, and persist its load balancing policies.

{\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
not enforced by the runtime, the system allows strings to be injected through
the admin daemon, constructing the Lua script in C++ is clunky, and the
administrator can inject really bad policies (e.g., while 1) that brings the
whole system down. Malacology exposes the important subsystems of Ceph's
logical storage device framework so that the MDS can safely and dynamically
load policies specified as Lua classes.

@@

\section{Evaluation} \label{evaluation} We have conducted our evaluation on
CloudLab as well as our own 20 node cluster. We follow the ``Popper
Convention"~\cite{anonymous} so as to try and make our experiments as
reproducible as possible. We have a GitHub repository with our source code,
experimental environment, configuration parameters, and workloads but due to
the double-blind policy, we have temporarily disabled the links to our graphs
and experimental details.

Much of this work focuses on the programmability of Ceph and the services that
can be layered on top of Malacology. As such, we focus on the performance of
the components and subsystems exposed by the framework. Detailed system-wide
performance measurements and analysis is forthcoming for both Mantle and ZLog,
as we try to find the best metadata balancing policies and physical design
implementation trade-offs respectively.

\subsection{ZLog}

We evaluate the mapping of ZLog onto Malacology by examining the costs of
utilizing the metadata service as a sequencer, and propagating interface
changes throughout the cluster using the monitoring service. We have omitted
evaluation of the overhead costs of adding object interfaces themselves due to
lack of space, and the fact that they are heavily tested and used already in
production with acceptable performance, introducing a small constant overhead
per operation but otherwise does not affect scalability.

@@

\subsection{MDS as a ZLog Sequencer}

The following experiments demonstrate the feasibility of using the Ceph MDS
cluster as a sequencer. The goal of the workloads is to saturate the MDS
cluster, so we design sequencers uses a more heavy-weight file system
operation. The experiments use our internal cluster and there are 15 clients
and between 1 and 3 MDS nodes.

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-seq-thruput.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] A
ZLog sequencer that uses the CephFS MDS enjoys the load balancing capabilities
of the metadata cluster. The ``Sequencer Baseline'' has no background
sequencers, while the ``2 Seq'' curves are the throughput of 1 sequencer when
an additional sequencer is running in the background generating
load.}\label{fig:mantle-seq-thruput}
\end{figure}

\subsubsection{Load Balancing}

We use a balancer policy specifically designed for the sequencer. The balancer
aggressively sheds half its load to its neighbor as soon as it can.  It is
based off the ``Greedy Spill Even" approach in the
Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced here for the reader's
convenience.

@@

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-cpu-utilization-annotate.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput} migrates to a new
MDS immediately because its uses a policy designed for bursty workloads. This
is a trace of the same workload running twice: once with a single MDS and once
with Mantle.}
\label{fig:mantle-cpu-utilization}
\end{figure}

Figure~\ref{fig:mantle-seq-thruput} gives a macro-level view of the performance
of the Ceph components exposed by Malacology. Each curve shows the throughput
from the perspective of one sequencer (\(y\) axis) over time (\(x\) axis).  The
``Sequencer Baseline'' curve is one sequencer serving 7 clients.  The sequencer
operates on an MDS cluster where balancing is turned off and there are no other
workloads running. ``No Load Balancing" adds another sequencer of equal load
running in the background. The performance of the background sequencer is
omitted. The last two policies, ``Sequencer Balancing Policy" and
``Conservative Balancing Policy", show the performance of the sequencer (with a
background sequencer running) when using Mantle with a 3 MDS node cluster.
These policies differ by the threshold at which they start balancing:
``Sequencer Balancing Policy" starts balancing immediately and ``Conservative
Balancing Policy" waits for 120 seconds.

Figure~\ref{fig:mantle-cpu-utilization} gives a time series graph to show the
behavior of the same workload, 2 sequencers, with balancing turned on and off.
The curves are the throughput (\(y\) axis) over time (\(x\) axis) and the graph
has been partitioned into the two cluster setups. The CPU utilization is
confined to one core the MDS is not multi-threaded, as it has been shown that
metadata performance does not improve with multiple
threads~\cite{konstantinos:pdsw2014-lustre-metadata}.

Overall, the clients achieve lower performance than the capability experiments
in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
operation to manage the capability.  ``Sequencer Baseline" is the best
performance for this particular workload because there is no contention.
Performance is stable and the MDS is under-loaded because we do not see the
spike at the end like the other curves. It appears the bottleneck is somewhere
with the clients -- perhaps it is our 1GBit/s network link. ``2 Seq; No Load
Balancing" shows an overloaded sequencer because the performance of our one
sequencer is low and unstable (high variation in performance). We also deduce
that it is overloaded by comparing to the ``2 Seq; Mantle Load Balancing"
curve, which shows improved performance when load is migrated. At 10 seconds,
the sequencer balancer immediately starts balancing and decides that two
sequencer streams should be split onto two MDSs. It migrates to its neighbor as
per the policy.

\textbf{The Malacology Approach:} Re-use the Ceph metadata migration, load
balancing, cache coherence, and capabilities subsystems of the MDS. Applying
Mantle's load balancing policy shows 2.02\(\times\) performance improvement and
16\% degradation in performance over the baseline, even though the baseline has
half as many clients.

>>>>

\item An implementation of Mantle, the programmable load balancing
service~\cite{sevilla:sc15-mantle}

@@

\begin{table*}
\centering
\title{Hi}
\begin{tabular}{  l | l | l | l }
\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Versioning}
  & Zookeeper coordination~\cite{hunt_zookeeper_2010}, Chubby~\cite{website:chubby}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols~\cite{CITEME}
  & serialization \& batching
  \\
{\it Logical Device}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
Beacon
  & Apache RabbitMQ~\cite{CITEME}
  & inter-daemon messages, centralized logging
  & debugging \& monitoring
  \\
\end{tabular}
\caption{The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces contributed in this paper.}
\label{table:examples}
\end{table*}

@@

{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
metadata load balancer, has been implemented on Malacology so that its
infrastructure can enjoy all the properties of existing internal abstractions. The load
balancing logic is intact but the infrastructure has been
improved to safely load, manage, and persist its load balancing policies.

{\bf Safely Loading Classes.} The original Mantle API is fragile: the API is
not enforced by the runtime, the system allows strings to be injected through
the admin daemon, constructing the Lua script in C++ is clunky, and the
administrator can inject really bad policies (e.g., while 1) that brings the
whole system down. Using the Malacology durability and beacon interfaces,
Mantle can safely load policies from the object store and centralize
warnings/errors, respectively.

@@

\section{Evaluation}
\label{evaluation}

We wanted to implement CORFU and we had two choices: build it from the ground
up or do it on top of Malacology. To show the feasability of building on
Malacology our evaluation focuses on the performance of the internal
abstractions that support the Malacogy interfaces. First, we benchmark
scenarios where contention is at the sequencer by examining the intefaces used
to map ZLog onto Malacology; specifically, we describe the sequencer
implementation and the propogation logical device interfaces. Next, we
benchmark scenarios when there are multiple logs by using Mantle to (1)
balance sequencers across a cluster and (2) recover from failure.

Detailed system-wide performance measurements and analysis is forthcoming for
both Mantle and ZLog, as we try to find the best metadata balancing policies
and physical design implementation trade-offs respectively.  Since this work
focuses on the programmabilty of Malacology, the goal of the benchmarks is to
show that the components and subsystems that support the Malacology interfaces
provide reasonable performance.

We have conducted our evaluation on CloudLab as well as our own 20 node
cluster. We follow the ``Popper Convention"~\cite{anonymous} so as to try and
make our experiments as reproducible as possible. We have a GitHub repository
with our source code, experimental environment, configuration parameters, and
workloads but due to the double-blind policy, we have temporarily disabled the
links to our graphs and experimental details.

\subsection{Mapping ZLog onto Malacology}
\label{mapping-zlog-onto-malacology}
We evaluate the mapping of ZLog onto Malacology by examining the costs of
utilizing the metadata service as a sequencer, and propagating interface
changes throughout the cluster using the monitoring service. We have omitted
evaluation of the overhead costs of adding object interfaces themselves due to
lack of space, and the fact that they are heavily tested and used already in
production with acceptable performance, introducing a small constant overhead
per operation but otherwise does not affect scalability.

\subsubsection{Sequencer Implementation}

@@

\subsection{ZLog and Mantle in a Multi-tenant Environment}

For scenarios with multiple shared logs and many clients, co-locating
sequencers on the same physical node is not ideal. Unfortunately, building a
load balancer that can migrate the shared resource (in this case, the resource
that guards the current token) would require building subsystems for migrating
resources, monitoring the workloads, monitoring physical nodes, partitioning
resources, and managing multiple sequencers.

The following experiments demonstrate the feasibility of using the mechanisms
of the Malacology load balancing interface (e.g., migrating, monitoring, and
partitioning load) to (1) alleviate load when multiple sequencers overwhelm a
physical servers and (2) recover from failure.

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-seq-thruput.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] A
ZLog sequencer that uses the Malacology load balancing interface enjoys the
load balancing features of the of the metadata cluster. The ``Sequencer
Baseline'' has no background sequencers, while the rest of the curves show the
throughput of 1 sequencer when an additional sequencer is running in the
background generating load.}\label{fig:mantle-seq-thruput}
\end{figure}

\subsubsection{Load Balancing Sequencers}

We evaluate the load balancing interface by showing the performance
implications of shedding load. The experiments use our internal cluster which
uses 10 nodes to store objects, one node to monitor the cluster, and 3 nodes
that can accomodate sequencers. The workload is two logs, one with 7 clients
and the other with 8 clients; we chose this workload because this load is the
point at which a single physical server gets overloaded.

We use a balancer policy specifically designed for multiple sequencers. The
balancer aggressively sheds half its load to its neighbor as soon as it can.
It is based off the ``Greedy Spill Even" approach in the
Mantle~\cite{sevilla:sc15-mantle} paper but is reproduced in
Figure~\ref{listing:greedy-spill} for the reader's convenience.

@@

\begin{figure}[t!]
\centering
\includegraphics{figures/mantle-cpu-utilization-annotate.png}
\caption{[\href{https://github.com/double-blind-submitter/osdi16}{source}] The
sequencer using Mantle in Figure~\ref{fig:mantle-seq-thruput} migrates to a new
server immediately because it uses a policy designed for bursty workloads. This
is a trace of the same workload running twice: once with a single host server and once
with Mantle.}
\label{fig:mantle-cpu-utilization}
\end{figure}

Figure~\ref{fig:mantle-seq-thruput} shows a performance profile of one of the
sequencers as the load balancing interface migrates sequencers around the
cluster to alleviate load.  Each curve shows the throughput (\(y\) axis) over
time (\(x\) axis).  ``Sequencer Baseline'' is the performance of one sequencer
on one physical server serving 7 clients - there are no other sequencers or
logs in the system and the curve represents the maximum throughput of a
sequencer on an underutilized physical server.  ``No Load Balancing" adds a
second sequencer of equal load to the same physical server with the intent of
overloading the single physical server; recall that the throughput for this
second sequencer is omitted. Compared to the baseline, the overloaded physical
server with 2 sequencers has more than a 2\(\times\) throughput degradation.

The last two curves of Figure~\ref{fig:mantle-seq-thruput} show the performance
improvements when sequencers are load balanced across the cluster. Both curves
are the throughput of one of the sequencers when a background sequencer is
running. These policies differ by the threshold at which they start balancing:
``Sequencer Balancing Policy" starts balancing immediately and ``Conservative
Balancing Policy" waits for 120 seconds before migrating sequencers.

The variation of the sequencer throughputs helps us determine which sequencers
are on overloaded physical servers.  ``Sequencer Baseline" is the best
performance for this particular workload because there is no contention.
Performance is stable and the MDS is under-loaded because we do not see the
spike at the end like the other curves.  ``No Load Balancing" shows an
overloaded sequencer because the performance of the monitored sequencer is low
and unstable (high variation in performance). We also deduce that it is
overloaded by comparing to the ``Sequencer Balancing Policy" and ``Conservative
Balancing Policy" curves, which show improved performance when load is
migrated.

%Overall, the clients achieve lower performance than the capability experiments
%#in Figure~\ref{fig:captp} because the sequencer uses a slower metadata
%operation to manage the capability.  It appears the bottleneck is somewhere
%with the clients -- perhaps it is our 1GBit/s network link.

Figure~\ref{fig:mantle-cpu-utilization} gives a time series graph to show the
behavior of the same workload, 2 sequencers, with balancing turned on and off.
The curves are the throughput (\(y\) axis) over time (\(x\) axis) and the graph
has been partitioned into the two cluster setup stages. The CPU utilization is
confined to one core because load balancer interface is not multi-threaded.
Recall the the load balancing interface re-uses the POSIX metadata cluster
migration mechanisms and as it has been shown that metadata performance does
not improve with multiple threads~\cite{konstantinos:pdsw2014-lustre-metadata}.

\textbf{The Malacology Approach:} Re-use the Ceph metadata migration, load
balancing, cache coherence, and capabilities subsystems of the metadata
cluster. Applying Mantle's load balancing policy shows 2.02\(\times\)
performance improvement and only a 16\% degradation in performance over the
baseline, even though the baseline has half as many clients.
